---
title: "gene activity score calculation"
output: gene activity score matrix (gene*cell)
date: "2024-10-01"
editor: Xiaoying
---

Calculate gene active score input: peak_count_matrix: a peak_count matrix from scATAC-seq with peak \* cell organism: species type human: GRCh38 / mouse: GRCm38 output: a gene \* cell matrix, the elements represent the gene activity for the cell

```{r setup, include=FALSE}
library(Seurat)
library(ggplot2)
library(sctransform)
library(MAESTRO, lib.loc = "/usr/lib64/R/library")
library(reticulate)
library(qs)
atac_matrix <- qread("/bmbl_data/xiaoying/Example_Data.qsave")
use_python("~/.local/share/r-miniconda/envs/r-reticulate/bin/python", required = TRUE)
pbmc.gene <- ATACCalculateGenescore(atac_matrix)
print(pbmc.gene)
#qsave(pbmc.gene,"GAS_SF10592.qsave")
```

## R Markdown

Calculate peak-gene regulatory score input: peak_count_matrix: a peak_count matrix from scATAC-seq with peak \* cell organism: species type human: GRCh38 / mouse: GRCm38 output: a gene \* peak matrix, the elements represent the regulatory potential for peak to gene

```{r cars}
CalGenePeakScore <-
  function(peak_count_matrix, organism = "GRCh38") {
    pbmc_peak <- peak_count_matrix
    n <- nrow(pbmc_peak)
    dia <- diag(n)
    rownames(dia) <- rownames(pbmc_peak)
    colnames(dia) <- 1:ncol(dia)
    gene_peak <-
      ATACCalculateGenescore(dia,
                             organism = organism,
                             decaydistance = 10000,
                             model = "Enhanced")
    colnames(gene_peak) <- rownames(peak_count_matrix)
    return (gene_peak)
  }

peak_gene_reg <- CalGenePeakScore(atac_matrix, "GRCh38")
print(peak_gene_reg)
```
