---
title: "Gene module enrichment analysis"
author: "YOUR NAME"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r}
library(tidyverse)
library(Seurat)
library(GSVA)

```

# Load gene list and Seurat object data

Next, define the gene list and load the Seurat object that contains your scRNA-seq data.


```{r}

example_gene_list <- c("Gpx3",  "Glrx",   "Lbp",   "Cryab")

example_obj <- read_rds("combined.rds")

```

# Calculate enrichment using AddModuleScore function from the Seurat package

Use Seurat’s AddModuleScore() function to calculate enrichment scores for the gene module across cells in the dataset.


```{r}

DefaultAssay(example_obj) <- "RNA"
example_obj <- AddModuleScore(example_obj, features = list(example_gene_list=example_gene_list), name = "example_gene_list")

```
# Calculate enrichment using ssGSEA and gsva method from the GSVA package

The GSVA package provides two methods for enrichment analysis: ssGSEA and GSVA. We will calculate both enrichment scores for the gene module.



https://www.bioconductor.org/packages/release/bioc/vignettes/GSVA/inst/doc/GSVA.html 

```{r}

exp_data <- LayerData(example_obj, layer = "data")

# ssGSEA method
ssgseaPar <- ssgseaParam(exp_data, list(example_gene_list=example_gene_list))
ssgsea.score <- gsva(ssgseaPar, verbose=T)
example_obj@meta.data[, 'example_ssgsea_score'] <- as.numeric(ssgsea.score)

# gsva method
gsvaPar <- ssgseaParam(exp_data, list(example_gene_list=example_gene_list))
gsva.score <- gsva(gsvaPar, verbose=T)
example_obj@meta.data[, 'example_gsva_score'] <- as.numeric(gsva.score)


```

# Gene set score visualization

## Violin plot

visualize the module and enrichment scores across cell types using violin plots.


```{r, fig.width=9, fig.height=6}

VlnPlot(example_obj, "example_gene_list1", group.by = "celltype") + ggtitle(paste0("example_gene_list - Module score"))

VlnPlot(example_obj, "example_ssgsea_score", group.by = "celltype") + ggtitle(paste0("example_gene_list - SSGSEA score"))

VlnPlot(example_obj, "example_gsva_score", group.by = "celltype") + ggtitle(paste0("example_gene_list - GSVA score"))

```


## Heatmap

visualize the individual genes in the module across cell types using heatmap

```{r, fig.width=6, fig.height=6}

example_obj <- ScaleData(example_obj)
heatmap_matrix <- AverageExpression(example_obj[example_gene_list,], assays = "RNA", slot = "scale.data")$RNA

pheatmap::pheatmap(
  heatmap_matrix,
  #annotation_col = my_sample_col,
  #scale = "row",
  color = colorRampPalette(c(
    "#344298", "#bbe0ed", "#fdf9b7", "#fba95e", "#ae0825"
  ))(1000),
  cluster_rows = T,
  cluster_cols = F,
  angle_col = 45,
  cellheight = 16,
  cellwidth = 16,
  main = "Average expression of example gene module"
)
  
```


```{r}

```


```{r}

```


```{r}

```



