---
title: "Pathway enrichment analysis and visualization using clusterProfiler"
author: "YOUR NAME"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This R Markdown tutorial was created to perform enrichment analysis on Gene Ontology. GO terms are organized in a directed acyclic graph, where edges between terms represent parent-child relationship.

If you use clusterProfiler in published research, please cite:

G Yu, LG Wang, Y Han, QY He. clusterProfiler: an R package for comparing biological themes among gene clusters. OMICS: A Journal of Integrative Biology 2012, 16(5):284-287. doi:[10.1089/omi.2011.0118](http://dx.doi.org/10.1089/omi.2011.0118)



# Load packages

```{r}
library(tidyverse)
library(clusterProfiler)
library(ggplot2)
library(org.Hs.eg.db)
```

# Load gene list and convert them to EntrezID

```{r}

example_gene_list <- c("GPX3",  "GLRX",   "LBP",   "CRYAB", "DEFB1", "HCLS1",   "SOD2",   "HSPA2",
       "ORM1",  "IGFBP1", "PTHLH", "GPC3",  "IGFBP3","TOB1",    "MITF",   "NDRG1",
       "NR1H4", "FGFR3",  "PVR",   "IL6",   "PTPRM", "ERBB2",   "NID2",   "LAMB1",
       "COMP",  "PLS3",   "MCAM",  "SPP1",  "LAMC1", "COL4A2",  "COL4A1", "MYOC",
       "ANXA4", "TFPI2",  "CST6",  "SLPI",  "TIMP2", "CPM",     "GGT1",   "NNMT",
       "MAL",   "EEF1A2", "HGD",   "TCN2",  "CDA",   "PCCA",    "CRYM",   "PDXK",
       "STC1",  "WARS",  "HMOX1", "FXYD2", "RBP4",   "SLC6A12", "KDELR3", "ITM2B")

example_gene_mapping = bitr(example_gene_list, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")

example_gene_id <- example_gene_mapping$ENTREZID

# GeneList is a pre-defined background gene list, and can be replaced with all genes in your datasets
data(geneList, package="DOSE")

```

# GO over-representation analysis

The clusterProfiler package implements enrichGO() for gene ontology over-representation test.

Here we use the dataset geneList provided by DOSE.

Gene Ontology defines concepts/classes used to describe gene function, and relationships between these concepts. It classifies functions along three aspects:

- MF: Molecular Function: molecular activities of gene products

- CC: Cellular Component: where gene products are active

- BP: Biological Process: pathways and larger processes made up of the activities of multiple gene products

GO terms are organized in a directed acyclic graph, where edges between terms represent parent-child relationship.


```{r}
gobp_result <- enrichGO(gene  = example_gene_id,
                universe      = names(geneList),
                OrgDb         = org.Hs.eg.db,
                ont           = "BP", # can be either one: BP, MF, CC
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
        readable      = TRUE)
head(gobp_result)
```

# Reactome pathway over-representation analysis

Enrichment analysis is a widely used approach to identify biological themes. ReactomePA implemented enrichPathway() that uses hypergeometric model to assess whether the number of selected genes associated with a reactome pathway is larger than expected.



```{r}
#BiocManager::install("ReactomePA")
library(ReactomePA)
reactome_result <- enrichPathway(gene=example_gene_id, pvalueCutoff = 0.05, readable=TRUE)
head(reactome_result)

```
# Pathway visualization

## Bar plot

```{r, fig.width=10, fig.height=10}
mutate(gobp_result, qscore = -log(p.adjust, base=10)) %>% 
    barplot(x="qscore")

```


## Dot plot

```{r, fig.width=10, fig.height=10}

dotplot(reactome_result, showCategory=30) + ggtitle("dotplot for result")
```
## Gene-Concept Network

```{r, fig.width=10, fig.height=10}
p1 <- cnetplot(reactome_result, node_label="category", 
        cex_label_category = 1.2) 
p2 <- cnetplot(reactome_result, node_label="gene", 
        cex_label_gene = 0.8) 
p3 <- cnetplot(reactome_result, node_label="all") 
p4 <- cnetplot(reactome_result, node_label="none", 
        color_category='firebrick', 
        color_gene='steelblue') 
cowplot::plot_grid(p1, p2, p3, p4, ncol=2, labels=LETTERS[1:4])
```

## Enrichment Map

```{r, fig.width=10, fig.height=10}
library(enrichplot)
edo <- pairwise_termsim(reactome_result)
p1 <- emapplot(edo)
p2 <- emapplot(edo, cex_category=1.5) # If using different version of enrichplot, may need to change `cex_category=` to `size_category=`
p3 <- emapplot(edo, layout="kk")
p4 <- emapplot(edo, cex_category=1.5,layout="kk") # If using different version of enrichplot, may need to change `cex_category=` to `size_category=`
cowplot::plot_grid(p1, p2, p3, p4, ncol=2, labels=LETTERS[1:4])
```


```{r}

```


```{r}

```


```{r}

```



