---
title: "4. Cell type annotation"
author: "Cankun Wang"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(error = F)
knitr::opts_chunk$set(fig.width = 8)

library(Seurat)
library(cowplot)
library(ggplot2)
library(patchwork)
library(here)
library(qs)
library(Polychrome)
library(RColorBrewer)
library(tidyverse)

```


# Set working dir

```{r,,eval=T,message=F,warning=F,error=F}
here::i_am("4_cell_type_annotation.Rmd")
here::set_here()
print(paste("Current working directory:", here::here()))

```

# Load data

```{r}
source("functions.R")
dir.create('../sample_obj', showWarnings = F)
combined <- qs::qread('combined.qsave')

DefaultAssay(combined) <- "RNA"
Idents(combined) <- combined$orig.ident

sample_color <-
  as.character(glasbey.colors()[-1])

cell_type_color <-
  as.character(palette36.colors(36)[-2])
two_color <- c('#C0C0C0', '#B00D23')

provided_marker <- read_csv("../marker/marker_by_day.csv") %>%
  mutate(cell_type = factor(cell_type))

provided_marker
```


```{r, fig.width=10, fig.height=8}

DimPlot(
  combined,
  reduction = "umap",
  cols = cell_type_color,
  label = T,
  pt.size = 0.4,
  repel = T,
  label.box = T
)

```

# Annotate by each sample

## Con0

Con0: all iPSCs

```{r, fig.height = 6}
# Settings
this_sample_name <- "Con0"
this_day <- "day0"

this_markers <- provided_marker %>%
  filter(str_detect(day, this_day)) %>%
  pull(gene)

Idents(combined) <- combined$orig.ident

# Load pre-computed object if exist
obj_filename <- paste0("../sample_obj/", this_sample_name, ".qsave")
if(file.exists(obj_filename)) {
  this_combined <- qs::qread(obj_filename)
} else {
  this_combined <- subset(combined, idents = this_sample_name)
  this_combined <- FindVariableFeatures(this_combined, verbose = F)
  this_combined <- RunPCA(this_combined, verbose = F)
  this_combined <- RunUMAP(this_combined, dims = 1:20, verbose = F)
  this_combined <- FindNeighbors(this_combined, verbose = F)
  this_combined <- FindClusters(this_combined, resolution = 0.2)
}


Idents(this_combined) <- this_combined$seurat_clusters
DimPlot(this_combined, reduction = "umap", label = T, label.box = T)

FeaturePlot(
  this_combined,
  features = unique(this_markers),
  ncol = 1,
  keep.scale = NULL,
  pt.size = 0.6,
  cols = two_color
)

```


```{r, fig.height=6}

tmp_ident <- as.factor(this_combined$seurat_clusters)
levels(tmp_ident) <- rep("iPSCs", length(levels(tmp_ident)))
this_combined <- AddMetaData(this_combined, tmp_ident, col.name = "cell_type")


Idents(this_combined) <- this_combined$cell_type
DimPlot(this_combined, reduction = "umap")

VlnPlot(this_combined,
        features = unique(this_markers),
        ncol = 1)
# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, ".qsave"))

table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

```
  

## N1KO0

N1KO0: all iPSCs

```{r, fig.height = 6}
# Settings
this_sample_name <- "N1KO0"
this_day <- "day0"

this_markers <- provided_marker %>%
  filter(str_detect(day, this_day)) %>%
  pull(gene)

Idents(combined) <- combined$orig.ident

# Load pre-computed object if exist
obj_filename <- paste0("../sample_obj/", this_sample_name, ".qsave")
if (file.exists(obj_filename)) {
  this_combined <- qs::qread(obj_filename)
} else {
  this_combined <- subset(combined, idents = this_sample_name)
  this_combined <- FindVariableFeatures(this_combined, verbose = F)
  this_combined <- RunPCA(this_combined, verbose = F)
  this_combined <- RunUMAP(this_combined, dims = 1:20, verbose = F)
  this_combined <- FindNeighbors(this_combined, verbose = F)
  this_combined <- FindClusters(this_combined, resolution = 0.2)
}


Idents(this_combined) <- this_combined$seurat_clusters
DimPlot(this_combined, reduction = "umap", label = T, label.box = T)

FeaturePlot(
  this_combined,
  features = unique(this_markers),
  ncol = 1,
  keep.scale = NULL,
  pt.size = 0.6,
  cols = two_color
)

```


```{r, fig.height=6}



tmp_ident <- as.factor(this_combined$seurat_clusters)
levels(tmp_ident) <- rep("iPSCs", length(levels(tmp_ident)))
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")


Idents(this_combined) <- this_combined$cell_type
DimPlot(this_combined, reduction = "umap")

VlnPlot(this_combined,
        features = unique(this_markers),
        ncol = 1)
# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, ".qsave"))

table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

```


## Con2

Con2: iPSCs, Mesoderm, Cardiac mesoderm

```{r, fig.height = 6}
# Settings
this_sample_name <- "Con2"
this_day <- "day2"

this_markers <- provided_marker %>%
  filter(str_detect(day, this_day)) %>%
  pull(gene)

Idents(combined) <- combined$orig.ident

# Load pre-computed object if exist
obj_filename <- paste0("../sample_obj/", this_sample_name, ".qsave")
if (file.exists(obj_filename)) {
  this_combined <- qs::qread(obj_filename)
  
} else {
  this_combined <- subset(combined, idents = this_sample_name)
  this_combined <- RunPCA(this_combined, verbose = F)
  this_combined <- RunUMAP(this_combined, dims = 1:20, verbose = F)
  this_combined <- FindNeighbors(this_combined, verbose = F)
  this_combined <-
    FindClusters(this_combined, resolution = 2.6, verbose = F)
}


Idents(this_combined) <- this_combined$seurat_clusters
DimPlot(this_combined, reduction = "umap", label = T, label.box = T)


```

```{r, fig.height = 6, fig.width=10}

FeaturePlot(
  this_combined,
  features = unique(this_markers),
  ncol = 3,
  keep.scale = NULL,
  pt.size = 0.6,
  cols = two_color
)

#VlnPlot(this_combined, features = this_markers, ncol = 3)

```


```{r, fig.height=6}


tmp_ident <- as.factor(this_combined$seurat_clusters)
tmp_levels <- rep("Mesoderm", 19)
tmp_levels[c(8, 11, 14, 16, 17, 19)] <- "iPSCs"
tmp_levels[c(7, 15)] <- "Cardiac mesoderm"

levels(tmp_ident) <- tmp_levels


this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")


Idents(this_combined) <- this_combined$cell_type
DimPlot(this_combined, reduction = "umap")

VlnPlot(this_combined,
        features = unique(this_markers),
        ncol = 3)

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, ".qsave"))

table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))


```
  

## N1KO2

N1KO2: iPSCs, Mesoderm, Cardiac mesoderm

```{r, fig.height = 6}
# Settings
this_sample_name <- "N1KO2"
this_day <- "day2"

this_markers <- provided_marker %>%
  filter(str_detect(day, this_day)) %>%
  pull(gene)

Idents(combined) <- combined$orig.ident

# Load pre-computed object if exist
obj_filename <- paste0("../sample_obj/", this_sample_name, ".qsave")
if (file.exists(obj_filename)) {
  this_combined <- qs::qread(obj_filename)
} else {
  this_combined <- subset(combined, idents = this_sample_name)
  this_combined <- RunPCA(this_combined, verbose = F)
  this_combined <- RunUMAP(this_combined, dims = 1:20, verbose = F)
  this_combined <- FindNeighbors(this_combined, verbose = F)
  this_combined <-
    FindClusters(this_combined, resolution = 0.3, verbose = F)
}


Idents(this_combined) <- this_combined$seurat_clusters
DimPlot(this_combined, reduction = "umap", label = T, label.box = T)

```

```{r, fig.height = 6, fig.width = 10}

FeaturePlot(
  this_combined,
  features = unique(this_markers),
  ncol = 3,
  keep.scale = NULL,
  pt.size = 0.6,
  cols = two_color
)

#VlnPlot(this_combined, features = this_markers, ncol = 3)

```


```{r, fig.height=6}

tmp_ident <- as.factor(this_combined$seurat_clusters)

tmp_levels <- rep("Mesoderm", 4)
tmp_levels[c(4)] <- "iPSCs"
#tmp_levels[c(3)] <- "Cardiac mesoderm"

levels(tmp_ident) <- tmp_levels

this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")


Idents(this_combined) <- this_combined$cell_type
DimPlot(this_combined, reduction = "umap")

VlnPlot(this_combined,
        features = unique(this_markers),
        ncol = 3)

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, ".qsave"))

table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

```


## Con5

Con5: iPSCs, Mesoderm, Cardiac mesoderm, SHF progenitor, FHF progenitor, Epicardial progenitor

```{r, fig.height = 6}
# Settings
this_sample_name <- "Con5"
this_day <- "day5"

this_markers <- provided_marker %>%
  filter(str_detect(day, this_day)) %>%
  pull(gene)

Idents(combined) <- combined$orig.ident

# Load pre-computed object if exist
obj_filename <- paste0("../sample_obj/", this_sample_name, ".qsave")
if (file.exists(obj_filename)) {
  this_combined <- qs::qread(obj_filename)
} else {
  this_combined <- subset(combined, idents = this_sample_name)
  this_combined <- FindVariableFeatures(this_combined, verbose = F)
  this_combined <- RunPCA(this_combined, verbose = F)
  this_combined <- RunUMAP(this_combined, dims = 1:20, verbose = F)
  this_combined <- FindNeighbors(this_combined, verbose = F)
  this_combined <-
    FindClusters(this_combined, resolution = 1.2, verbose = F)
}


Idents(this_combined) <- this_combined$seurat_clusters
DimPlot(this_combined, reduction = "umap", label = T, label.box = T)

```

```{r, fig.width=10, fig.height = 8}

FeaturePlot(
  this_combined,
  features = unique(this_markers),
  ncol = 3,
  keep.scale = NULL,
  pt.size = 0.6,
  cols = two_color
)


```



```{r, fig.height=6}

deg_filename <- paste0("../sample_obj/", this_sample_name, "_marker.csv")
if (!file.exists(deg_filename)) {
  cts_markers <-
    FindAllMarkers(this_combined,
                   only.pos = T,
                   min.pct = 0.5,
                   logfc.threshold = 0.3) %>%
    filter(pct.1 > 0.25)
  write.csv(cts_markers,
            paste0("../sample_obj/", this_sample_name, "_marker.csv"))
}

tmp_ident <- as.factor(this_combined$seurat_clusters)

tmp_levels <- rep("Cardiac mesoderm", length(levels(tmp_ident)))
tmp_levels[c(3, 6, 8, 9)] <- "Differentiating iPSCs"
tmp_levels[c(4)] <- "SHF progenitor"
levels(tmp_ident) <- tmp_levels

this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")


Idents(this_combined) <- this_combined$cell_type
DimPlot(this_combined, reduction = "umap")

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, ".qsave"))
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

```

```{r, fig.width=10, fig.height = 12}

VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```



## N1KO5

N1KO5: Differentiating iPSCs, Mesoderm, Cardiac mesoderm, SHF progenitor, FHF progenitor, Epicardial progenitor

```{r, fig.height = 6}
# Settings
this_sample_name <- "N1KO5"
this_day <- "day5"

this_markers <- provided_marker %>%
  filter(str_detect(day, this_day)) %>%
  pull(gene)

Idents(combined) <- combined$orig.ident

# Load pre-computed object if exist
obj_filename <- paste0("../sample_obj/", this_sample_name, ".qsave")
if(file.exists(obj_filename)) {
  this_combined <- qs::qread(obj_filename)
} else {
  this_combined <- subset(combined, idents = this_sample_name)
  this_combined <- FindVariableFeatures(this_combined, verbose = F)
  this_combined <- RunPCA(this_combined, verbose = F)
  this_combined <- RunUMAP(this_combined, dims = 1:20, verbose = F)
  this_combined <- FindNeighbors(this_combined, verbose = F)
  this_combined <- FindClusters(this_combined, resolution = 2, verbose = F)
}


Idents(this_combined) <- this_combined$seurat_clusters
DimPlot(this_combined, reduction = "umap", label = T, label.box = T)


```

```{r, fig.width=10, fig.height = 8}

FeaturePlot(
  this_combined,
  features = unique(this_markers),
  ncol = 3,
  keep.scale = NULL,
  pt.size = 0.6,
  cols = two_color
)
#VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```


```{r, fig.height=6}

deg_filename <- paste0("../sample_obj/", this_sample_name, "_marker.csv")
if (!file.exists(deg_filename)) {
  cts_markers <-
    FindAllMarkers(this_combined,
                   only.pos = T,
                   min.pct = 0.5,
                   logfc.threshold = 0.3)
  write.csv(cts_markers,
            paste0("../sample_obj/", this_sample_name, "_marker.csv"))
}


tmp_ident <- as.factor(this_combined$seurat_clusters)
tmp_levels <- rep("Cardiac mesoderm", length(levels(tmp_ident)))
tmp_levels[c(14)] <- "Differentiating iPSCs"
tmp_levels[c(12, 13)] <- "SHF progenitor"
levels(tmp_ident) <- tmp_levels


this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")


Idents(this_combined) <- this_combined$cell_type
DimPlot(this_combined, reduction = "umap")

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, ".qsave"))
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

```

```{r, fig.width=10, fig.height = 12}

VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```



## Con10
 
Con10: Cardiac mesoderm, SHF progenitor, FHF progenitor, Epicardial progenitor, Atrial cardiomyocytes, Ventricular cardiomyocytes, Pacemaker cells, Vascular smooth muscle cells, Cardiac fibroblast, Early cardiomyocytes

```{r, fig.height = 6}
# Settings
this_sample_name <- "Con10"
this_day <- "day10"

this_markers <- provided_marker %>%
  filter(str_detect(day, this_day)) %>%
  pull(gene)

Idents(combined) <- combined$orig.ident

# Load pre-computed object if exist
obj_filename <- paste0("../sample_obj/", this_sample_name, ".qsave")
if(file.exists(obj_filename)) {
  this_combined <- qs::qread(obj_filename)
} else {
  this_combined <- subset(combined, idents = this_sample_name)
  this_combined <- FindVariableFeatures(this_combined, verbose = F)
  this_combined <- RunPCA(this_combined, verbose = F)
  this_combined <- RunUMAP(this_combined, dims = 1:20, verbose = F)
  this_combined <- FindNeighbors(this_combined, verbose = F)
  this_combined <- FindClusters(this_combined, resolution = 0.8, verbose = F)
}


Idents(this_combined) <- this_combined$seurat_clusters
DimPlot(this_combined, reduction = "umap", label = T, label.box = T)


```

```{r, fig.width=10, fig.height = 16}

FeaturePlot(
  this_combined,
  features = unique(this_markers),
  ncol = 3,
  keep.scale = NULL,
  pt.size = 0.8,
  cols = two_color
)

#VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```


```{r, fig.height=6}

deg_filename <- paste0("../sample_obj/", this_sample_name, "_marker.csv")
if (!file.exists(deg_filename)) {
  cts_markers <-
    FindAllMarkers(this_combined,
                   only.pos = T,
                   min.pct = 0.5,
                   logfc.threshold = 0.3)
  write.csv(cts_markers,
            paste0("../sample_obj/", this_sample_name, "_marker.csv"))
}

tmp_ident <- as.factor(this_combined$seurat_clusters)
tmp_levels <- rep("Early cardiomyocytes", length(levels(tmp_ident)))
tmp_levels[c(9, 11, 14)] <- "FHF progenitor"
tmp_levels[c(3, 5)] <- "SHF progenitor"
tmp_levels[c(6)] <- "Epicardial progenitor"
tmp_levels[c(1, 7, 8, 12, 13)] <- "Unidentified"
levels(tmp_ident) <- tmp_levels

this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")


Idents(this_combined) <- this_combined$cell_type
DimPlot(this_combined, reduction = "umap")

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, ".qsave"))

#Con10
table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

```

```{r, fig.width=10, fig.height = 20}

VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```


## N1KO10
 
N1KO10: Cardiac mesoderm, SHF progenitor, FHF progenitor, Epicardial progenitor, Atrial cardiomyocytes, Ventricular cardiomyocytes, Pacemaker cells, Vascular smooth muscle cells, Cardiac fibroblast, Early cardiomyocytes

```{r, fig.height = 6}
# Settings
this_sample_name <- "N1KO10"
this_day <- "day10"

this_markers <- provided_marker %>%
  filter(str_detect(day, this_day)) %>%
  pull(gene)

Idents(combined) <- combined$orig.ident

# Load pre-computed object if exist
obj_filename <- paste0("../sample_obj/", this_sample_name, ".qsave")
if(file.exists(obj_filename)) {
  this_combined <- qs::qread(obj_filename)
} else {
  this_combined <- subset(combined, idents = this_sample_name)
  this_combined <- FindVariableFeatures(this_combined, verbose = F)
  this_combined <- RunPCA(this_combined, verbose = F)
  this_combined <- RunUMAP(this_combined, dims = 1:20, verbose = F)
  this_combined <- FindNeighbors(this_combined, verbose = F)
  this_combined <- FindClusters(this_combined, resolution = 1, verbose = F)
}


Idents(this_combined) <- this_combined$seurat_clusters
DimPlot(this_combined, reduction = "umap", label = T, label.box = T)



```

```{r, fig.width=10, fig.height = 16}

FeaturePlot(
  this_combined,
  features = unique(this_markers),
  ncol = 3,
  keep.scale = NULL,
  pt.size = 0.6,
  cols = two_color
)

#VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```


```{r, fig.height=6}

deg_filename <- paste0("../sample_obj/", this_sample_name, "_marker.csv")
if (!file.exists(deg_filename)) {
  cts_markers <-
    FindAllMarkers(this_combined,
                   only.pos = T,
                   min.pct = 0.5,
                   logfc.threshold = 0.3)
  write.csv(cts_markers,
            paste0("../sample_obj/", this_sample_name, "_marker.csv"))
}

tmp_ident <- as.factor(this_combined$seurat_clusters)
tmp_levels <- rep("Epicardial progenitor", length(levels(tmp_ident)))
tmp_levels[c(12, 13)] <- "FHF progenitor"
tmp_levels[c(2, 7, 8)] <- "SHF progenitor"
tmp_levels[c(3, 5, 9, 14)] <- "Early cardiomyocytes"




levels(tmp_ident) <- tmp_levels


this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")


Idents(this_combined) <- this_combined$cell_type
DimPlot(this_combined, reduction = "umap")

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, ".qsave"))

#N1KO10
table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

```

```{r, fig.width=10, fig.height = 20}

VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```



## Con14
 
Con14: SHF progenitor, FHF progenitor, Epicardial progenitor, Atrial cardiomyocytes, Ventricular cardiomyocytes, Pacemaker cells, Vascular smooth muscle cells, Cardiac fibroblast


```{r, fig.height = 6}
# Settings
this_sample_name <- "Con14"
this_day <- "day14"

this_markers <- provided_marker %>%
  filter(str_detect(day, this_day)) %>%
  pull(gene)

Idents(combined) <- combined$orig.ident

# Load pre-computed object if exist
obj_filename <- paste0("../sample_obj/", this_sample_name, ".qsave")
if(file.exists(obj_filename)) {
  this_combined <- qs::qread(obj_filename)
} else {
  this_combined <- subset(combined, idents = this_sample_name)
  this_combined <- FindVariableFeatures(this_combined, verbose = F)
  this_combined <- RunPCA(this_combined, verbose = F)
  this_combined <- RunUMAP(this_combined, dims = 1:20, verbose = F)
  this_combined <- FindNeighbors(this_combined, verbose = F)
  this_combined <- FindClusters(this_combined, resolution = 1, verbose = F)
}


Idents(this_combined) <- this_combined$seurat_clusters
DimPlot(this_combined, reduction = "umap", label = T, label.box = T)



```

```{r, fig.width=10, fig.height = 12}

FeaturePlot(
  this_combined,
  features = unique(this_markers),
  ncol = 3,
  keep.scale = NULL,
  pt.size = 0.6,
  cols = two_color
)
#VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```


```{r, fig.height=6}

deg_filename <- paste0("../sample_obj/", this_sample_name, "_marker.csv")
if (!file.exists(deg_filename)) {
  cts_markers <-
    FindAllMarkers(this_combined,
                   only.pos = T,
                   min.pct = 0.5,
                   logfc.threshold = 0.3)
  write.csv(cts_markers,
            paste0("../sample_obj/", this_sample_name, "_marker.csv"))
}

tmp_ident <- as.factor(this_combined$seurat_clusters)
tmp_levels <- rep("Early cardiomyocytes", length(levels(tmp_ident)))
tmp_levels[c(1, 7, 9)] <- "Cardiac fibroblast"
tmp_levels[c(3, 8, 11, 13 )] <- "Unidentified"
levels(tmp_ident) <- tmp_levels

this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")

Idents(this_combined) <- this_combined$cell_type
DimPlot(this_combined, reduction = "umap")

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, ".qsave"))

table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

```

```{r, fig.width=10, fig.height = 20}

VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```



## N1KO14
 
N1KO14: SHF progenitor, FHF progenitor, Epicardial progenitor, Atrial cardiomyocytes, Ventricular cardiomyocytes, Pacemaker cells, Vascular smooth muscle cells, Cardiac fibroblast


```{r, fig.height = 6}
# Settings
this_sample_name <- "N1KO14"
this_day <- "day14"

this_markers <- provided_marker %>%
  filter(str_detect(day, this_day)) %>%
  pull(gene)

Idents(combined) <- combined$orig.ident

# Load pre-computed object if exist
obj_filename <- paste0("../sample_obj/", this_sample_name, ".qsave")
if(file.exists(obj_filename)) {
  this_combined <- qs::qread(obj_filename)
} else {
  this_combined <- subset(combined, idents = this_sample_name)
  this_combined <- FindVariableFeatures(this_combined, verbose = F)
  this_combined <- RunPCA(this_combined, verbose = F)
  this_combined <- RunUMAP(this_combined, dims = 1:20, verbose = F)
  this_combined <- FindNeighbors(this_combined, verbose = F)
  this_combined <- FindClusters(this_combined, resolution = 1, verbose = F)
}


Idents(this_combined) <- this_combined$seurat_clusters
DimPlot(this_combined, reduction = "umap", label = T, label.box = T)



```

```{r, fig.width=10, fig.height = 12}

FeaturePlot(
  this_combined,
  features = unique(this_markers),
  ncol = 3,
  keep.scale = NULL,
  pt.size = 0.6,
  cols = two_color
)

#VlnPlot(this_combined, features = unique(this_markers), ncol = 3)


```


```{r, fig.height=6}

deg_filename <- paste0("../sample_obj/", this_sample_name, "_marker.csv")
if (!file.exists(deg_filename)) {
  cts_markers <-
    FindAllMarkers(this_combined,
                   only.pos = T,
                   min.pct = 0.5,
                   logfc.threshold = 0.3)
  write.csv(cts_markers,
            paste0("../sample_obj/", this_sample_name, "_marker.csv"))
}

tmp_ident <- as.factor(this_combined$seurat_clusters)
tmp_levels <- rep("Early cardiomyocytes", length(levels(tmp_ident)))
tmp_levels[c(5, 8, 12, 16)] <- "Cardiac fibroblast"
levels(tmp_ident) <- tmp_levels



this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")


Idents(this_combined) <- this_combined$cell_type
DimPlot(this_combined, reduction = "umap")

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, ".qsave"))

table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

```

```{r, fig.width=10, fig.height = 20}

VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```



## Con30
 
Con30: Atrial cardiomyocytes, Ventricular cardiomyocytes, Pacemaker cells, Vascular smooth muscle cells, Cardiac fibroblast


```{r, fig.height = 6}
# Settings
this_sample_name <- "Con30"
this_day <- "day30"

this_markers <- provided_marker %>%
  filter(str_detect(day, this_day)) %>%
  pull(gene)

Idents(combined) <- combined$orig.ident

# Load pre-computed object if exist
obj_filename <- paste0("../sample_obj/", this_sample_name, ".qsave")
if(file.exists(obj_filename)) {
  this_combined <- qs::qread(obj_filename)
} else {
  this_combined <- subset(combined, idents = this_sample_name)
  this_combined <- FindVariableFeatures(this_combined, verbose = F)
  this_combined <- RunPCA(this_combined, verbose = F)
  this_combined <- RunUMAP(this_combined, dims = 1:20, verbose = F)
  this_combined <- FindNeighbors(this_combined, verbose = F)
  this_combined <- FindClusters(this_combined, resolution = 2.6, verbose = F)
}


Idents(this_combined) <- this_combined$seurat_clusters
DimPlot(this_combined, reduction = "umap", label = T, label.box = T)



```

```{r, fig.width=10, fig.height = 12}

FeaturePlot(
  this_combined,
  features = unique(this_markers),
  ncol = 3,
  keep.scale = NULL,
  pt.size = 0.6,
  cols = two_color
)


#VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```


```{r, fig.height=6}

deg_filename <- paste0("../sample_obj/", this_sample_name, "_marker.csv")
if (!file.exists(deg_filename)) {
  cts_markers <-
    FindAllMarkers(this_combined,
                   only.pos = T,
                   min.pct = 0.5,
                   logfc.threshold = 0.3)
  write.csv(cts_markers,
            paste0("../sample_obj/", this_sample_name, "_marker.csv"))
}

tmp_ident <- as.factor(this_combined$seurat_clusters)
tmp_levels <- rep("Ventricular cardiomyocytes", length(levels(tmp_ident)))
tmp_levels[c(4)] <- "Atrial cardiomyocytes"
tmp_levels[c(11)] <- "Pacemaker cell"
tmp_levels[c(12)] <- "Cardiac fibroblast"

levels(tmp_ident) <- tmp_levels


this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")


Idents(this_combined) <- this_combined$cell_type
DimPlot(this_combined, reduction = "umap")

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, ".qsave"))

table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

```

```{r, fig.width=10, fig.height = 16}

VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```


## N1KO30
 
N1KO30: Atrial cardiomyocytes, Ventricular cardiomyocytes, Pacemaker cells, Vascular smooth muscle cells, Cardiac fibroblast

```{r, fig.height = 6}
# Settings
this_sample_name <- "N1KO30"
this_day <- "day30"

this_markers <- provided_marker %>%
  filter(str_detect(day, this_day)) %>%
  pull(gene)

Idents(combined) <- combined$orig.ident

# Load pre-computed object if exist
obj_filename <- paste0("../sample_obj/", this_sample_name, ".qsave")
if (file.exists(obj_filename)) {
  this_combined <- qs::qread(obj_filename)
} else {
  this_combined <- subset(combined, idents = this_sample_name)
  this_combined <- FindVariableFeatures(this_combined, verbose = F)
  this_combined <- RunPCA(this_combined, verbose = F)
  this_combined <- RunUMAP(this_combined, dims = 1:20, verbose = F)
  this_combined <- FindNeighbors(this_combined, verbose = F)
  this_combined <-
    FindClusters(this_combined, resolution = 2, verbose = F)
}


Idents(this_combined) <- this_combined$seurat_clusters
DimPlot(this_combined, reduction = "umap", label = T, label.box = T)



```

```{r, fig.width=10, fig.height = 12}

FeaturePlot(
  this_combined,
  features = unique(this_markers),
  ncol = 3,
  keep.scale = NULL,
  pt.size = 0.6,
  cols = two_color
)
#VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```


```{r, fig.height=6}

deg_filename <- paste0("../sample_obj/", this_sample_name, "_marker.csv")
if (!file.exists(deg_filename)) {
  cts_markers <-
    FindAllMarkers(this_combined,
                   only.pos = T,
                   min.pct = 0.5,
                   logfc.threshold = 0.3)
  write.csv(cts_markers,
            paste0("../sample_obj/", this_sample_name, "_marker.csv"))
}

tmp_ident <- as.factor(this_combined$seurat_clusters)
tmp_levels <- rep("Atrial cardiomyocytes", length(levels(tmp_ident)))
tmp_levels[c(8, 12, 13, 14)] <- "Ventricular cardiomyocytes"
tmp_levels[c(11)] <- "Pacemaker cell"
tmp_levels[c(2)] <- "Cardiac fibroblast"
tmp_levels[c(10)] <- "Vascular smooth muscle cells"

levels(tmp_ident) <- tmp_levels



this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")


Idents(this_combined) <- this_combined$cell_type
DimPlot(this_combined, reduction = "umap")

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, ".qsave"))

table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

```

```{r, fig.width=10, fig.height = 20}

VlnPlot(this_combined, features = unique(this_markers), ncol = 3)

```


# UMAP of cell types using all samples
  
```{r, eval=T, fig.width=10, fig.height=6}
obj1 <- qread("../sample_obj/Con0.qsave")
obj2 <- qread("../sample_obj/Con2.qsave")
obj3 <- qread("../sample_obj/Con5.qsave")
obj4 <- qread("../sample_obj/Con10.qsave")
obj5 <- qread("../sample_obj/Con14.qsave")
obj6 <- qread("../sample_obj/Con30.qsave")
obj7 <- qread("../sample_obj/N1KO0.qsave")
obj8 <- qread("../sample_obj/N1KO2.qsave")
obj9 <- qread("../sample_obj/N1KO5.qsave")
obj10 <- qread("../sample_obj/N1KO10.qsave")
obj11 <- qread("../sample_obj/N1KO14.qsave")
obj12 <- qread("../sample_obj/N1KO30.qsave")


obj <-
  merge(
    x = obj1,
    y = c(obj2,
          obj3,
          obj4,
          obj5,
          obj6,
          obj7,
          obj8,
          obj9,
          obj10,
          obj11,
          obj12)
  )

combined <- AddMetaData(combined, obj$cell_type, col.name = "cell_type")

tmp_ident <- as.factor(combined$orig.ident)
levels(tmp_ident) <- c(rep("Con",6), rep("KO",6))
combined <- AddMetaData(combined, tmp_ident, col.name = "condition")

```


```{r fig.width=10, fig.height=6}


Idents(combined) <- combined$cell_type
DimPlot(combined,
        reduction = "umap",
        cols = cell_type_color,
        pt.size = 0.8)

```


```{r fig.width=10, fig.height=4}

Idents(combined) <- combined$cell_type

p1 <- DimPlot(
  combined,
  reduction = "umap",
  split.by = "condition",
  cols = cell_type_color,
  pt.size = 0.8,
  ncol = 6
)

print(p1)

```


```{r fig.width=10, fig.height=4}

Idents(combined) <- combined$cell_type

p1 <- DimPlot(
  combined,
  reduction = "umap",
  split.by = "orig.ident",
  cols = cell_type_color,
  pt.size = 0.8,
  ncol = 6
)

print(p1)

```


```{r,eval=T,message=T,warning=F,error=F}
qs::qsave(combined,'combined.qsave')
```

