---
title: "Cell type annotation for project 2 by sample"
author: "Cankun Wang"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: true
    toc_float: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(error = F)
knitr::opts_chunk$set(fig.width = 8)

library(Seurat)
library(cowplot)
library(ggplot2)
library(patchwork)
library(here)
library(qs)
library(Polychrome)
library(RColorBrewer)
library(tidyverse)

```


# Set working dir

```{r,,eval=T,message=F,warning=F,error=F}
here::i_am("cell_type_annotation_by_sample.rmd")
here::set_here()
print(paste("Current working directory:", here::here()))
set.seed(42)
```

# Load data

```{r}

source("../shared/functions.R")
dir.create('../annotation_obj', showWarnings = F)


provided_marker <- read_csv("../marker/marker_by_day.csv") %>%
  mutate(cell_type = factor(cell_type))

provided_marker


```

```{r, eval=T,echo=F}
print_total_pct <- function(cat1, cat2) {
  total_cells <- as.numeric(colSums(table(cat1, cat2)))

  total_pct <-
    matrix(nrow = nrow(
      table(cat1, cat2)
    ))
  pct_cells <- for (i in 1:length(total_cells)) {
    this_sample <-
      colnames(table(cat1, cat2))[i]
    this_pct <-
      as.data.frame(round(table(cat1, cat2)[, i] / total_cells[i], 3))  # added round function here
    colnames(this_pct) <- this_sample
    total_pct <- cbind(total_pct, this_pct)
  }
  
  total_pct <- total_pct[, -1]
  return (total_pct)
}


```


# Annotate by each sample

Control Day10, Control Day14,  Control Day5,  Control Day8,      HG Day10,      HG Day14,       HG Day5,       HG Day8 

Figure order:

1. UMAP colored by clusters and sample

2. Provided marker gene expression in UMAP

3. Provided marker gene expression in violin plots

4. Provided marker gene expression in dot plots

5. Top 5 marker gene expression for each cluster in dot plots

## Day5 - Control

```{r day5wt, fig.height=8, fig.width=10}

this_day <- "day5"
this_sample_name <- "Control Day5"

this_combined <- qs::qread(paste0("../sample_obj/", this_sample_name, "_processed.qsave"))
Idents(this_combined) <- this_combined$seurat_clusters

p1 <- DimPlot(
  this_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  label = T,
  pt.size = 0.5
)+ theme(legend.position = "top")

print(p1)


```


```{r, fig.height=7, fig.width=10, eval=T}

# Cell type with marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type_with_marker")

Idents(this_combined) <- this_combined$cell_type_with_marker

this_combined <- RenameIdents(this_combined, 
                              "3" = "Cardiac mesoderm (PDGFRA+, MESP1+)",
                              "4" = "Epithelial-like cell (EPCAM+, KRT18+)",
                              "6" = "Mesoderm (POU5F1+, TBXT1+)",
                              "0" = "Ectoderm progenitor (LSAMP1+)",
                              "1" = "Mesoderm (POU5F1+, TBXT1+)",
                              "2" = "Mesoderm (POU5F1+, TBXT1+)",
                              "5" = "Ectoderm progenitor (LSAMP1+)",
                              "7" = "Neuron progenitor (DCC+, EYA2+)"
                              )

DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type_with_marker")

# Cell type without marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")

Idents(this_combined) <- this_combined$cell_type

this_combined <- RenameIdents(this_combined, 
                              "3" = "Cardiac mesoderm",
                              "4" = "Epithelial-like cell",
                              "6" = "Mesoderm",
                              "0" = "Ectoderm progenitor",
                              "1" = "Mesoderm",
                              "2" = "Mesoderm",
                              "5" = "Ectoderm progenitor",
                              "7" = "Neuron progenitor"
                              )

DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type")



# percentage of cell types on all samples 
table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, "_cell_type.qsave"))

```


## Day5 - HG

```{r day5h, fig.height=8, fig.width=10}

this_day <- "day5"
this_sample_name <- "HG Day5"
this_combined <- qs::qread(paste0("../sample_obj/", this_sample_name, "_processed.qsave"))
Idents(this_combined) <- this_combined$seurat_clusters

p1 <- DimPlot(
  this_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  label = T,
  pt.size = 0.5
)+ theme(legend.position = "top")

print(p1)


```


```{r, fig.height=7, fig.width=10, eval=T}

# Cell type with marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type_with_marker")

Idents(this_combined) <- this_combined$cell_type_with_marker

this_combined <- RenameIdents(this_combined, 
                              "0" = "Neuron progenitor (MAP2+, VLDLR-AS1+)",
                              "1" = "Mesoderm (POU5F1+)",
                              "2" = "Mesoderm (POU5F1+)",
                              "3" = "Cardiac mesoderm (POU5F1+, PDGFRA+)",
                              "4" = "Ectoderm progenitor (PAX6+)"
                              )

DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type_with_marker")

# Cell type without marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")

Idents(this_combined) <- this_combined$cell_type

this_combined <- RenameIdents(this_combined, 
                              "0" = "Neuron progenitor",
                              "1" = "Mesoderm",
                              "2" = "Mesoderm",
                              "3" = "Cardiac mesoderm",
                              "4" = "Ectoderm progenitor"
                              )

DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type")



# percentage of cell types on all samples 
table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, "_cell_type.qsave"))

```





## Day8 - Control

```{r day8wt, fig.height=8, fig.width=10}

this_day <- "day8"
this_sample_name <- "Control Day8"
this_combined <- qs::qread(paste0("../sample_obj/", this_sample_name, "_processed.qsave"))
Idents(this_combined) <- this_combined$seurat_clusters

p1 <- DimPlot(
  this_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  label = T,
  pt.size = 0.5
)+ theme(legend.position = "top")

print(p1)

```


```{r, fig.height=7, fig.width=10, eval=T}

# Cell type with marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type_with_marker")

Idents(this_combined) <- this_combined$cell_type_with_marker


this_combined <- RenameIdents(this_combined, 
                              "3" = "Neural crest cell (PAX3+)",
                              "4" = "Endocardial cell (KDR+, NFATc1+, PECAM1+, CDH5+)",
                              "6" = "Neuron cell (NEUROD4+, NFASC+, ELAVL3+)",
                              "0" = "FHF progenitor (NKX2-5+, HAND1+, TBX5+, ISL1-)",
                              "1" = "Cardiac mesoderm (PDGFRA+)",
                              "2" = "SHF progenitor (NKX2-5+, ISL1+, HAND2+)",
                              "5" = "Endocardial cushion cell (NPR3+, CDH11+)"
                              )

DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type_with_marker")

# Cell type without marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")

Idents(this_combined) <- this_combined$cell_type


this_combined <- RenameIdents(this_combined, 
                              "3" = "Neural crest cell",
                              "4" = "Endocardial cell",
                              "6" = "Neuron cell",
                              "0" = "FHF progenitor",
                              "1" = "Cardiac mesoderm",
                              "2" = "SHF progenitor",
                              "5" = "Endocardial cushion cell"
                              )

DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type")



# percentage of cell types on all samples 
table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, "_cell_type.qsave"))



```



## Day8 - HG

```{r day8h, fig.height=8, fig.width=10}

this_day <- "day8"
this_sample_name <- "HG Day8"
this_combined <- qs::qread(paste0("../sample_obj/", this_sample_name, "_processed.qsave"))
Idents(this_combined) <- this_combined$seurat_clusters

p1 <- DimPlot(
  this_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  label = T,
  pt.size = 0.5
)+ theme(legend.position = "top")

print(p1)


```


```{r, fig.height=7, fig.width=10, eval=T}
# Cell type with marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type_with_marker")

Idents(this_combined) <- this_combined$cell_type_with_marker

this_combined <- RenameIdents(this_combined, 
                              "0" = "Cardiac mesoderm (PDGFRA+)",
                              "1" = "Neuronal progenitor (FGF13+,)",
                              "2" = "Neuronal progenitor (RMST+, LMX1A+, NAV3+, UNC5C+)",
                              "3" = "Proliferative neuronal progenitor (CENPF+, UBE2C+, TOP2A+, CCNB1+)",
                              "4" = "Neural crest cell (SOX10+, TFAP2A+, PAX3+)",
                              "5" = "Neural crest derived cell (CRABP1+, STMN2+, SNCG+, TUBB3+)",
                              "6" = "Cardiac progenitor (TNNT2+)",
                              "7" = "Epithelial cell (EPCAM+, CLDN3,6,7+, KRT8, KRT18+, KRT19+)",
                              "8" = "Endothelial cell (CDH5+, PECAM1+, KDR+)"
                              )

DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type_with_marker")

# Cell type without marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")

Idents(this_combined) <- this_combined$cell_type

this_combined <- RenameIdents(this_combined, 
                              "0" = "Cardiac mesoderm",
                              "1" = "Neuronal progenitor",
                              "2" = "Neuronal progenitor",
                              "3" = "Proliferative neuronal progenitor",
                              "4" = "Neural crest cell",
                              "5" = "Neural crest derived cell",
                              "6" = "Cardiac progenito",
                              "7" = "Epithelial cell",
                              "8" = "Endothelial cell"
                              )


DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type")



# percentage of cell types on all samples 
table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, "_cell_type.qsave"))


```






## Day10 - Control

```{r day10wt, fig.height=8, fig.width=10}

this_day <- "day10"
this_sample_name <- "Control Day10"
this_combined <- qs::qread(paste0("../sample_obj/", this_sample_name, "_processed.qsave"))
Idents(this_combined) <- this_combined$seurat_clusters

p1 <- DimPlot(
  this_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  label = T,
  pt.size = 0.5
)+ theme(legend.position = "top")

print(p1)


```


```{r, fig.height=7, fig.width=10, eval=T}
# Cell type with marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type_with_marker")

Idents(this_combined) <- this_combined$cell_type_with_marker

this_combined <- RenameIdents(this_combined, 
                              "0" = "Early cardiomyocyte (ACTC1, MYH6+, TNNT2+, TNNC1+, MYL3+, and MYL4+)",
                              "1" = "Endocardial cushion cell (NTRK2+, CDH11+)",
                              "2" = "FHF-derived cardiomyocyte and SHF-derived cardiomyocyte (TNNT2+, TBX5+, HAND1+, HAND2+)",
                              "3" = "FHF-derived cardiomyocyte and SHF-derived cardiomyocyte (TNNT2+, TBX5+, HAND1+, HAND2+)",
                              "4" = "Neuron cell (LMX1A+, CNTNAP3B+, GDF7+)",
                              "5" = "SHF-derived cardiomyocyte (NKX2-5+, ISL1+, TNNT2+)",
                              "6" = "Neuron cell (ELAV4+, STMN2+, MYT1L+, EBF2+)",
                              "7" = "Unknown (expressing lipid metabolism genes)",
                              "8" = "Endocardial cushion cell (NPR3+)",
                              "9" = "Endothelial cell (CDH5+, KDR+)"
                              )


DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type_with_marker")

# Cell type without marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")

Idents(this_combined) <- this_combined$cell_type

this_combined <- RenameIdents(this_combined, 
                              "0" = "Early cardiomyocyte",
                              "1" = "Endocardial cushion cell",
                              "2" = "FHF-derived cardiomyocyte and SHF-derived cardiomyocyte",
                              "3" = "FHF-derived cardiomyocyte and SHF-derived cardiomyocyte",
                              "4" = "Neuron cell",
                              "5" = "SHF-derived cardiomyocyte",
                              "6" = "Neuron cell",
                              "7" = "Unknown",
                              "8" = "Endocardial cushion cell",
                              "9" = "Endothelial cell"
                              )


DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type")



# percentage of cell types on all samples 
table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, "_cell_type.qsave"))

```


## Day10 - HG

```{r day10h, fig.height=8, fig.width=10}

this_day <- "day10"
this_sample_name <- "HG Day10"
this_combined <- qs::qread(paste0("../sample_obj/", this_sample_name, "_processed.qsave"))
Idents(this_combined) <- this_combined$seurat_clusters

p1 <- DimPlot(
  this_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  label = T,
  pt.size = 0.5
)+ theme(legend.position = "top")

print(p1)

```



```{r, fig.height=7, fig.width=10, eval=T}
# Cell type with marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type_with_marker")

Idents(this_combined) <- this_combined$cell_type_with_marker

this_combined <- RenameIdents(this_combined, 
                              "0" = "Endocardial cushion cell (NTRK2+, CDH11+)",
                              "1" = "Valvular interstitial cell (NTRK2+, CDH11+, PPRX1+, S100A4+, ACTA2+)",
                              "2" = "Endocardial cushion cell (NTRK2+, CDH11+)",
                              "3" = "FHF & SHF progenitor (NKX2-5+, ISL1+, TBX5+)",
                              "4" = "Neural crest cell (SOX10+, PAX3+, TFAP2B+)",
                              "5" = "Epithelial-like cell (EPCAM+, KRT19+)",
                              "6" = "Endocardial cushion cell (CHD11+)",
                              "7" = "Neuron cell (DCC+, STMN2+, NEFM+)",
                              "8" = "Endothelial cell (KDR+, PECAM1+, CDH5+)"
                              )


DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type_with_marker")

# Cell type without marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")

Idents(this_combined) <- this_combined$cell_type

this_combined <- RenameIdents(this_combined, 
                              "0" = "Endocardial cushion cell",
                              "1" = "Valvular interstitial cell",
                              "2" = "Endocardial cushion cell",
                              "3" = "FHF & SHF progenitor",
                              "4" = "Neural crest cell",
                              "5" = "Epithelial-like cell",
                              "6" = "Endocardial cushion cell",
                              "7" = "Neuron cell",
                              "8" = "Endothelial cell"
                              )



DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type")



# percentage of cell types on all samples 
table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, "_cell_type.qsave"))

```




## Day14 - Control

```{r day14wt, fig.height=8, fig.width=10}

this_day <- "day14"
this_sample_name <- "Control Day14"
this_combined <- qs::qread(paste0("../sample_obj/", this_sample_name, "_processed.qsave"))
Idents(this_combined) <- this_combined$seurat_clusters

p1 <- DimPlot(
  this_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  label = T,
  pt.size = 0.5
)+ theme(legend.position = "top")

print(p1)

```


```{r, fig.height=7, fig.width=10, eval=T}
# Cell type with marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type_with_marker")

Idents(this_combined) <- this_combined$cell_type_with_marker
this_combined <- RenameIdents(this_combined, 
                              "0" = "Ventricular cardiomyocyte (MYL7+, MYH7+, TNNT2+, IRX4+, MYL2+)",
                              "1" = "Ventricular cardiomyocyte (MYL7+, MYH7+, TNNT2+, IRX4+, MYL2+)",
                              "2" = "Ventricular cardiomyocyte (MYL7+, MYH7+, TNNT2+, IRX4+, MYL2+)",
                              "3" = "Proliferating ventricular cardiomyocyte (IRX4+, TNNT2+, CENPF+, HMGB2+, TOP2A+, MKI67+)",
                              "4" = "Cardiac fibroblast (DDR2+)",
                              "5" = "Atrial cardiomyocyte (MYH6+, NR2F2+, TNNT2+)",
                              "6" = "Unknown (CDH11+)",
                              "7" = "Endothelial cell (PECAM1+, CDH5+)"
                              )



DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type_with_marker")

# Cell type without marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")

Idents(this_combined) <- this_combined$cell_type

this_combined <- RenameIdents(this_combined, 
                              "0" = "Ventricular cardiomyocyte",
                              "1" = "Ventricular cardiomyocyte",
                              "2" = "Ventricular cardiomyocyte",
                              "3" = "Proliferating ventricular cardiomyocyte",
                              "4" = "Cardiac fibroblast",
                              "5" = "Atrial cardiomyocyte",
                              "6" = "Unknown",
                              "7" = "Endothelial cell"
                              )


DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type")



# percentage of cell types on all samples 
table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, "_cell_type.qsave"))

```



## Day14 - HG

```{r day14h, fig.height=8, fig.width=10}

this_day <- "day14"
this_sample_name <- "HG Day14"
this_combined <- qs::qread(paste0("../sample_obj/", this_sample_name, "_processed.qsave"))
Idents(this_combined) <- this_combined$seurat_clusters

p1 <- DimPlot(
  this_combined,
  reduction = "umap",
  group.by = "seurat_clusters",
  label = T,
  pt.size = 0.5
)+ theme(legend.position = "top")

print(p1)

```

```{r, fig.height=7, fig.width=10, eval=T}
# Cell type with marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type_with_marker")

Idents(this_combined) <- this_combined$cell_type_with_marker

this_combined <- RenameIdents(this_combined, 
                              "0" = "Neuron cell (NPAS3+, RFX4+, CTNND2+)",
                              "1" = "Fibroblast (DDR2+, COL3A1+, COL1A1+)",
                              "2" = "Unknown (highly proliferative)",
                              "3" = "Neuronal cell (DPP10+, RMST+, CNTNAP2+, LMX1A+)",
                              "4" = "Neuron cell (STMN2+, DCC+, TUBB3+)",
                              "5" = "Ventricular cardiomyocyte (TNNT2+, MYH6+, MYH7+)",
                              "6" = "Neural crest cell (PAX3+, SOX10+)",
                              "7" = "Epithelial cell (KRT19+, EPCAM+)",
                              "8" = "Epithelial cell (EPCAM+ lipid metabolism genes)",
                              "9" = "Unknown (involved in cell fate determination)",
                              "10" = "Endothelial cell (KDR+, PECAM1+, CDH5+)"
                              )


DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type_with_marker")

# Cell type without marker name
tmp_ident <- as.factor(this_combined$seurat_clusters)
this_combined <-
  AddMetaData(this_combined, tmp_ident, col.name = "cell_type")

Idents(this_combined) <- this_combined$cell_type

this_combined <- RenameIdents(this_combined, 
                              "0" = "Neuron cell",
                              "1" = "Fibroblast",
                              "2" = "Unknown",
                              "3" = "Neuronal cell",
                              "4" = "Neuron cell",
                              "5" = "Ventricular cardiomyocyte",
                              "6" = "Neural crest cell",
                              "7" = "Epithelial cell",
                              "8" = "Epithelial cell",
                              "9" = "Unknown",
                              "10" = "Endothelial cell"
                              )

DimPlot(this_combined, reduction = "umap", label = T, pt.size = 0.7)

this_combined <-
  AddMetaData(this_combined, Idents(this_combined), col.name = "cell_type")



# percentage of cell types on all samples 
table(this_combined$cell_type)
table(this_combined$cell_type) / sum(as.numeric(table(this_combined$cell_type)))

# Save
qs::qsave(this_combined, paste0("../sample_obj/", this_sample_name, "_cell_type.qsave"))

```

# Merged results

```{r,eval=T,message=T,warning=F,error=F, fig.width=10, fig.height=8}

combined <- qs::qread("combined.qsave")
DimPlot(
  combined,
  reduction = "umap",
  label = T,
  pt.size = 0.4,
  repel = T,
  cols = cell_type_color,
  label.box = T
) + theme(legend.position = "none")


```

```{r,eval=T,message=T,warning=F,error=F, fig.width=12, fig.height=5}

named_sample_color <- sample_color[1:length(levels(as.factor(combined$sample)))]
names(named_sample_color) <- levels(as.factor(combined$sample))

Idents(combined) <- combined$sample
plots <- lapply(
  X = levels(combined),
  FUN = function(x) {
    #x = 'Healthy control iPSC-D5'
    this_color_index <- which(names(named_sample_color) == x)
    this_color <- as.character(named_sample_color[this_color_index])
    return(
      DimPlot(
        combined,
        reduction = "umap",
        pt.size = 0.6,
        cells.highlight = CellsByIdentities(combined, idents = x),
        cols.highlight = this_color
      ) +
        ggplot2::theme(legend.position = "top")
    )
  }
)

p2 <- CombinePlots(plots, ncol = 4)

p2

# png(
#   paste0("./umap_by_sample.png"),
#   width = 6800,
#   height = 3500,
#   res = 300
# )
# print(p2)
# dev.off()

```


```{r,eval=F,message=T,warning=F,error=F, fig.width=12, fig.height=5}

named_sample_color <- sample_color[1:length(levels(as.factor(combined$sample)))]
names(named_sample_color) <- levels(as.factor(combined$sample))

named_ct_color <- cell_type_color[1:length(levels(as.factor(combined$cell_type)))]
names(named_ct_color) <- levels(as.factor(combined$cell_type))

Idents(combined) <- combined$sample
plots <- lapply(
  X = levels(combined),
  FUN = function(x) {
    #x = 'Control Day5'
    Idents(combined) <- combined$sample
    this_combiend <- subset(combined, idents = x)
    this_color <- named_ct_color[unique(this_combiend$cell_type)]
    Idents(this_combiend) <- this_combiend$cell_type
    return(
      DimPlot(
        combined,
        reduction = "umap",
        group.by = "cell_type",
        pt.size = 0.5,
        cells.highlight = CellsByIdentities(this_combiend, idents = unique(this_combiend$cell_type)),
        cols.highlight = sort(this_color)
      ) +
        ggplot2::theme(legend.position = "right") + 
        ggtitle(x)
    )
  }
)

#print(plots[1])

p3 <- CombinePlots(plots, ncol = 2)

p3

 png(
   paste0("./umap_cell_type_by_sample.png"),
   width = 6800,
   height = 9500,
   res = 300
 )
 print(p3)
 dev.off()

```



```{r}
knitr::kable(table(combined$cell_type, combined$sample))


knitr::kable(print_total_pct(combined$cell_type, combined$sample))
```