---
title: "Data export for GEO submission"
author: "Cankun Wang"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(error = F)
knitr::opts_chunk$set(fig.width = 8)

library(Seurat)
library(cowplot)
library(ggplot2)
library(patchwork)
library(here)
library(qs)
library(Polychrome)
library(RColorBrewer)
library(tidyverse)
library(ComplexHeatmap)
library(data.table)
library(circlize)
library(RColorBrewer)
library(GSVA)
library(ggpubr)
library(matrixStats)
library(EnhancedVolcano)
```


# Set working dir

```{r,,eval=T,message=F,warning=F,error=F}

here::set_here()
print(paste("Current working directory:", here::here()))

```

# Load data

```{r}
source("functions.R")
dir.create('../sample_obj', showWarnings = F)
combined <- qs::qread('combined.qsave')

DefaultAssay(combined) <- "RNA"
Idents(combined) <- combined$orig.ident

sample_color <-
  as.character(glasbey.colors()[-1])

cell_type_color <-
  as.character(palette36.colors(36)[-2])
two_color <- c('#C0C0C0', '#B00D23')


provided_marker <- read_csv("../marker/marker_by_day.csv") %>%
  mutate(cell_type = factor(cell_type))

provided_marker

```


# Setup

```{r}
meta <- combined@meta.data
write.csv(meta, "metadata.csv", row.names = T, col.names = T, quote = F)
counts <- GetAssayData(combined, assay = "RNA", slot="counts")
write.csv(counts, "counts.csv", row.names = T, col.names = T, quote = F)


```

```{r, include=F}

```

# Heat by sample

```{r}

#this_combined <- qs::qread("../integration/day30_harmony.qsave")
this_combined <- combined

DefaultAssay(this_combined) <- "RNA"

this_combind <- ScaleData(this_combined, verbose = F)

this_genes <- unique(provided_marker$gene)

if (length(which(toupper(this_genes) == this_genes)) > length(this_genes) / 2) {
  this_genes_index <-
    which(toupper(rownames(this_combined)) %in% this_genes)
} else {
  this_genes_index <- which(rownames(this_combined) %in% this_genes)
}


mat <-
  AverageExpression(this_combind, slot = "data")$RNA[this_genes_index, ]
mat <- as.matrix(mat)

#mat <- log1p(mat)
mat <- (mat - as.numeric(rowMeans(mat))) / rowSds(mat)


color_sequence <- seq(min(mat), max(mat), length = 3) / 1
color_fun = colorRamp2(color_sequence, c("#3266AD", "white", "#B12424"), space = "RGB")

col_fun <- all_sample_cols
tmp_order <-
  c("Con0",
    "N1KO0",
    "Con2",
    "N1KO2",
    "Con5",
    "N1KO5",
    "Con10",
    "N1KO10",
    "Con14",
    "N1KO14",
    "Con30",
    "N1KO30")
names(col_fun) <- tmp_order
tmp_ident <- factor(this_combind$orig.ident, levels = tmp_order)

#mat <- mat[, c(4, 2, 3, 1)]
ht1 <- Heatmap(
  mat,
  top_annotation = HeatmapAnnotation(
    Group = tmp_order,
    col = list(Group = col_fun),
    show_annotation_name = F
  ),
  show_column_names = T,
  cluster_rows = T,
  show_row_names = T,
  row_km = NULL,
  show_column_dend = F,
  col = color_fun,
  heatmap_legend_param = list(
    title = "Relative\nexpression",
    at = c(min(mat), 0,  max(mat)),
    labels = c("low", "", "high")
  ),
  column_order = 1:ncol(mat),
  column_names_rot = 30
)



ht1 <- Heatmap(
  mat,
  show_column_names = T,
  cluster_rows = T,
  show_row_names = T,
  row_km = NULL,
  show_column_dend = F,
  col = color_fun,
  heatmap_legend_param = list(
    title = "Relative\nexpression",
    at = c(min(mat), 0,  max(mat)),
    labels = c("low", "", "high")
  ),
  column_order = 1:ncol(mat),
  column_names_rot = 30
)


draw(ht1)
```

# Heat by CT+sample

```{r}

#this_combined <- qs::qread("../integration/day30_harmony.qsave")
this_combined <- combined

DefaultAssay(this_combined) <- "RNA"
tmp_ident <- paste(this_combined$cell_type, this_combined$orig.ident,sep = "-")
this_combind <- AddMetaData(this_combind, tmp_ident, "ct_sample")
this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combind) <- this_combind$ct_sample
this_genes <- unique(provided_marker$gene)

if (length(which(toupper(this_genes) == this_genes)) > length(this_genes) / 2) {
  this_genes_index <-
    which(toupper(rownames(counts)) %in% this_genes)
} else {
  this_genes_index <- which(rownames(counts) %in% this_genes)
}


mat <-
  AverageExpression(this_combind, slot = "data")$RNA[this_genes_index, ]
mat <- as.matrix(mat)

#mat <- log1p(mat)
mat <- (mat - as.numeric(rowMeans(mat))) / rowSds(mat)


color_sequence <- seq(min(mat), max(mat), length = 3) / 2
color_fun = colorRamp2(color_sequence, c("#3266AD", "white", "#B12424"), space = "RGB")

#mat <- mat[, c(4, 2, 3, 1)]


ht1 <- Heatmap(
  mat,
  show_column_names = T,
  cluster_rows = T,
  show_row_names = T,
  row_km = NULL,
  show_column_dend = F,
  col = color_fun,
  heatmap_legend_param = list(
    title = "Relative\nexpression",
    at = c(min(mat), 0,  max(mat)),
    labels = c("low", "", "high")
  ),
  column_order = 1:ncol(mat),
  column_names_rot = 45
)

png("../figure/heatmap/marker_ct_sample.png", width = 5500, height = 2000, res = 300)

draw(ht1)
dev.off()
```

```{r}
Idents(combined) <- combined$orig.ident


#Mesoderm (d2) 
#
#Cardiac mesoderm (d5)
#
#cardiac progenitor (FHF, SHF, epicardial) (d10)
#
#early CM (D10)
#
#early CM (D14)
#
#atrial CM (d30)
#
#ventricular CM (d30)
#
#pacemaker (d30)

c1 <- subset(combined, idents = c("Con2", "N1KO2"))
Idents(c1) <- c1$cell_type
c1 <- subset(c1, idents = c("Mesoderm"))

c2 <- subset(combined, idents = c("Con5", "N1KO5"))
Idents(c2) <- c2$cell_type
c2 <- subset(c2, idents = c("Cardiac mesoderm"))

c3 <- subset(combined, idents = c("Con10", "N1KO10"))
Idents(c3) <- c3$cell_type
c3 <- subset(c3, idents = c("Unidentified"), invert = T)

c4 <- subset(combined, idents = c("Con14", "N1KO14"))
Idents(c4) <- c4$cell_type
c4 <- subset(c4, idents = c("Early cardiomyocytes"), invert = F)

c5 <- subset(combined, idents = c("Con30", "N1KO30"))
Idents(c5) <- c5$cell_type
c5 <- subset(c5, idents = c("Atrial cardiomyocytes","Pacemaker cell", "Ventricular cardiomyocytes"), invert = F)

selected_combined <- merge(c1, y=c(c2,c3,c4,c5))


tmp_ident <- paste0(selected_combined$cell_type, "-d",str_replace(selected_combined$orig.ident, "Con|N1KO",""))
table(tmp_ident)
selected_combined <- AddMetaData(selected_combined, tmp_ident, col.name = "ct_day")
Idents(selected_combined) <- selected_combined$ct_day
levels(Idents(selected_combined) )


```

# Heat by avg selected CT x markers

```{r}
this_combined <- selected_combined
DefaultAssay(this_combined) <- "RNA"
Idents(this_combined) <- this_combined$ct_day

#this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combind) <- this_combind$ct_day
this_genes <- provided_marker %>%
  dplyr::filter(!cell_type %in% c("iPSCs","Vascular smooth muscle cells","Cardiac fibroblast")) %>%
  dplyr::pull(gene) %>%
  unique()

if (length(which(toupper(this_genes) == this_genes)) > length(this_genes) / 2) {
  this_genes_index <-
    which(toupper(rownames(this_combined)) %in% this_genes)
} else {
  this_genes_index <- which(rownames(this_combined) %in% this_genes)
}


mat <-
  AverageExpression(this_combind, slot = "data")$RNA[this_genes_index, ]
mat <- as.matrix(mat)

#mat <- log1p(mat)
mat <- (mat - as.numeric(rowMeans(mat))) / rowSds(mat)


color_sequence <- seq(min(mat), max(mat), length = 3) / 1.5
color_fun = colorRamp2(color_sequence, c("#3266AD", "white", "#B12424"), space = "RGB")



mat <- mat[this_genes, ]

ht1 <- Heatmap(
  mat,
  show_column_names = T,
  cluster_rows = F,
  show_row_names = T,
  row_km = NULL,
  show_column_dend = F,
  col = color_fun,
  heatmap_legend_param = list(
    title = "Relative\nexpression",
    at = c(min(mat), 0,  max(mat)),
    labels = c("low", "", "high")
  ),
  column_order = 1:ncol(mat),
  column_names_rot = 30
)




pdf(
  paste0("../figure/heatmap_avg_selected_marker.pdf"),
  width = 12,
  height = 8
)

print(draw(ht1, padding = unit(c(0, 30, 0, 0), "mm")))
dev.off()


pdf(
  paste0("../figure/dotplot_avg_selected_marker.pdf"),
  width = 10,
  height = 6
)

DotPlot(this_combined, features = this_genes, cols = two_color) + theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()


```


# Heatmap by avg selected CT x markers (only WT)

```{r}
Idents(selected_combined) <- selected_combined$condition
this_combined <- subset(selected_combined, idents = "Con")
Idents(this_combined) <- this_combined$ct_day
DefaultAssay(this_combined) <- "RNA"

#this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combind) <- this_combind$ct_day
this_genes <- provided_marker %>%
  dplyr::filter(!cell_type %in% c("iPSCs","Vascular smooth muscle cells","Cardiac fibroblast")) %>%
  dplyr::pull(gene) %>%
  unique()

if (length(which(toupper(this_genes) == this_genes)) > length(this_genes) / 2) {
  this_genes_index <-
    which(toupper(rownames(this_combined)) %in% this_genes)
} else {
  this_genes_index <- which(rownames(this_combined) %in% this_genes)
}


mat <-
  AverageExpression(this_combind, slot = "data")$RNA[this_genes_index, ]
mat <- as.matrix(mat)

#mat <- log1p(mat)
mat <- (mat - as.numeric(rowMeans(mat))) / rowSds(mat)


color_sequence <- seq(min(mat), max(mat), length = 3) / 1.5
color_fun = colorRamp2(color_sequence, c("#3266AD", "white", "#B12424"), space = "RGB")



mat <- mat[this_genes, ]

ht1 <- Heatmap(
  mat,
  show_column_names = T,
  cluster_rows = F,
  show_row_names = T,
  row_km = NULL,
  show_column_dend = F,
  col = color_fun,
  heatmap_legend_param = list(
    title = "Relative\nexpression",
    at = c(min(mat), 0,  max(mat)),
    labels = c("low", "", "high")
  ),
  column_order = 1:ncol(mat),
  column_names_rot = 30
)




pdf(
  paste0("../figure/heatmap_avg_selected_marker_WT.pdf"),
  width = 12,
  height = 8
)

print(draw(ht1, padding = unit(c(0, 30, 0, 0), "mm")))
dev.off()


pdf(
  paste0("../figure/dotplot_avg_selected_marker_WT.pdf"),
  width = 10,
  height = 6
)

DotPlot(this_combined, features = this_genes, cols = two_color) + theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()

```

# Heatmap by avg selected CT x markers (only KO)

```{r}
Idents(selected_combined) <- selected_combined$condition
this_combined <- subset(selected_combined, idents = "KO")
Idents(this_combined) <- this_combined$ct_day
DefaultAssay(this_combined) <- "RNA"

#this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combind) <- this_combind$ct_day
this_genes <- provided_marker %>%
  dplyr::filter(!cell_type %in% c("iPSCs","Vascular smooth muscle cells","Cardiac fibroblast")) %>%
  dplyr::pull(gene) %>%
  unique()

if (length(which(toupper(this_genes) == this_genes)) > length(this_genes) / 2) {
  this_genes_index <-
    which(toupper(rownames(this_combined)) %in% this_genes)
} else {
  this_genes_index <- which(rownames(this_combined) %in% this_genes)
}


mat <-
  AverageExpression(this_combind, slot = "data")$RNA[this_genes_index, ]
mat <- as.matrix(mat)

#mat <- log1p(mat)
mat <- (mat - as.numeric(rowMeans(mat))) / rowSds(mat)


color_sequence <- seq(min(mat), max(mat), length = 3) / 1.5
color_fun = colorRamp2(color_sequence, c("#3266AD", "white", "#B12424"), space = "RGB")



mat <- mat[this_genes, ]

ht1 <- Heatmap(
  mat,
  show_column_names = T,
  cluster_rows = F,
  show_row_names = T,
  row_km = NULL,
  show_column_dend = F,
  col = color_fun,
  heatmap_legend_param = list(
    title = "Relative\nexpression",
    at = c(min(mat), 0,  max(mat)),
    labels = c("low", "", "high")
  ),
  column_order = 1:ncol(mat),
  column_names_rot = 30
)




pdf(
  paste0("../figure/heatmap_avg_selected_marker_KO.pdf"),
  width = 12,
  height = 8
)

print(draw(ht1, padding = unit(c(0, 30, 0, 0), "mm")))
dev.off()


pdf(
  paste0("../figure/dotplot_avg_selected_marker_KO.pdf"),
  width = 10,
  height = 6
)

DotPlot(this_combined, features = this_genes, cols = two_color) + theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()

```


# Heat by cells selected CT x markers

```{r}

this_combined <- selected_combined
DefaultAssay(this_combined) <- "RNA"
this_combined <- ScaleData(this_combined)


#this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combined) <- this_combined$ct_day
this_genes <-  provided_marker %>%
  dplyr::filter(!cell_type %in% c("iPSCs","Vascular smooth muscle cells","Cardiac fibroblast")) %>%
  dplyr::pull(gene) %>%
  unique()


mapal <- colorRampPalette(RColorBrewer::brewer.pal(6,"RdBu"))(128)

pdf(
  paste0("../figure/heatmap_cells_selected_marker.pdf"),
  width = 12,
  height = 6
)

print(DoHeatmap(this_combined, features = this_genes, angle = 90, label = F) + scale_fill_gradientn(colours = rev(mapal))) 
dev.off()

```

# Heat by cells selected CT x markers (only WT)

```{r}

Idents(selected_combined) <- selected_combined$condition
this_combined <- subset(selected_combined, idents = "Con")
Idents(this_combined) <- this_combined$ct_day
DefaultAssay(this_combined) <- "RNA"
this_combined <- ScaleData(this_combined)

#this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combined) <- this_combined$ct_day
this_genes <-  provided_marker %>%
  dplyr::filter(!cell_type %in% c("iPSCs","Vascular smooth muscle cells","Cardiac fibroblast")) %>%
  dplyr::pull(gene) %>%
  unique()


mapal <- colorRampPalette(RColorBrewer::brewer.pal(6,"RdBu"))(128)

pdf(
  paste0("../figure/heatmap_cells_selected_marker_WT.pdf"),
  width = 12,
  height = 6
)

print(DoHeatmap(this_combined, features = this_genes, angle = 90, label = F) + scale_fill_gradientn(colours = rev(mapal))) 
dev.off()

```


# Heat by cells selected CT x markers (only ko)

```{r}

Idents(selected_combined) <- selected_combined$condition
this_combined <- subset(selected_combined, idents = "KO")
Idents(this_combined) <- this_combined$ct_day
DefaultAssay(this_combined) <- "RNA"
this_combined <- ScaleData(this_combined)


#this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combined) <- this_combined$ct_day
this_genes <-  provided_marker %>%
  dplyr::filter(!cell_type %in% c("iPSCs","Vascular smooth muscle cells","Cardiac fibroblast")) %>%
  dplyr::pull(gene) %>%
  unique()


mapal <- colorRampPalette(RColorBrewer::brewer.pal(6,"RdBu"))(128)

pdf(
  paste0("../figure/heatmap_cells_selected_marker_KO.pdf"),
  width = 12,
  height = 6
)

print(DoHeatmap(this_combined, features = this_genes, angle = 90, label = F) + scale_fill_gradientn(colours = rev(mapal))) 
dev.off()

```

# Heatmap by top 3 markers 
```{r}
DefaultAssay(selected_combined) <- "RNA"
Idents(selected_combined) <- selected_combined$ct_day
#cts_markers <- FindAllMarkers(selected_combined, logfc.threshold = 0.8)
write.csv(cts_markers,"selected_combined_ct_day_marker.csv")
        
this_genes <- cts_markers %>%
  group_by(cluster) %>%
  arrange(cluster) %>%
  top_n(3, avg_log2FC) %>%
  pull(gene) %>%
  unique()

```
# Heat by avg top3 markers

```{r}
this_combined <- selected_combined
DefaultAssay(this_combined) <- "RNA"
Idents(this_combined) <- this_combined$ct_day

#this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combind) <- this_combind$ct_day

mat <-
  AverageExpression(this_combind, slot = "data")$RNA[this_genes, ]
mat <- as.matrix(mat)

#mat <- log1p(mat)
mat <- (mat - as.numeric(rowMeans(mat))) / rowSds(mat)


color_sequence <- seq(min(mat), max(mat), length = 3) / 1.5
color_fun = colorRamp2(color_sequence, c("#3266AD", "white", "#B12424"), space = "RGB")



mat <- mat[this_genes, ]

ht1 <- Heatmap(
  mat,
  show_column_names = T,
  cluster_rows = F,
  show_row_names = T,
  row_km = NULL,
  show_column_dend = F,
  col = color_fun,
  heatmap_legend_param = list(
    title = "Relative\nexpression",
    at = c(min(mat), 0,  max(mat)),
    labels = c("low", "", "high")
  ),
  column_order = 1:ncol(mat),
  column_names_rot = 30
)

pdf(
  paste0("../figure/heatmap_avg_top3.pdf"),
  width = 12,
  height = 8
)

print(draw(ht1, padding = unit(c(0, 30, 0, 0), "mm")))
dev.off()


pdf(
  paste0("../figure/dotplot_avg_top3.pdf"),
  width = 14,
  height = 6
)

DotPlot(this_combined, features = this_genes, cols = two_color) + theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()


```


# Heatmap by avg selected CT x markers (only WT)

```{r}
Idents(selected_combined) <- selected_combined$condition
this_combined <- subset(selected_combined, idents = "Con")
Idents(this_combined) <- this_combined$ct_day
DefaultAssay(this_combined) <- "RNA"

#this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combind) <- this_combind$ct_day

if (length(which(toupper(this_genes) == this_genes)) > length(this_genes) / 2) {
  this_genes_index <-
    which(toupper(rownames(this_combined)) %in% this_genes)
} else {
  this_genes_index <- which(rownames(this_combined) %in% this_genes)
}

mat <-
  AverageExpression(this_combind, slot = "data")$RNA[this_genes_index, ]
mat <- as.matrix(mat)

#mat <- log1p(mat)
mat <- (mat - as.numeric(rowMeans(mat))) / rowSds(mat)


color_sequence <- seq(min(mat), max(mat), length = 3) / 1.5
color_fun = colorRamp2(color_sequence, c("#3266AD", "white", "#B12424"), space = "RGB")



mat <- mat[this_genes, ]

ht1 <- Heatmap(
  mat,
  show_column_names = T,
  cluster_rows = F,
  show_row_names = T,
  row_km = NULL,
  show_column_dend = F,
  col = color_fun,
  heatmap_legend_param = list(
    title = "Relative\nexpression",
    at = c(min(mat), 0,  max(mat)),
    labels = c("low", "", "high")
  ),
  column_order = 1:ncol(mat),
  column_names_rot = 30
)




pdf(
  paste0("../figure/heatmap_avg_top3_WT.pdf"),
  width = 12,
  height = 8
)

print(draw(ht1, padding = unit(c(0, 30, 0, 0), "mm")))
dev.off()


pdf(
  paste0("../figure/dotplot_avg_top3_WT.pdf"),
  width = 14,
  height = 6
)

DotPlot(this_combined, features = this_genes, cols = two_color) + theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()

```

# Heatmap by avg selected CT x markers (only KO)

```{r}
Idents(selected_combined) <- selected_combined$condition
this_combined <- subset(selected_combined, idents = "KO")
Idents(this_combined) <- this_combined$ct_day
DefaultAssay(this_combined) <- "RNA"

#this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combind) <- this_combind$ct_day
if (length(which(toupper(this_genes) == this_genes)) > length(this_genes) / 2) {
  this_genes_index <-
    which(toupper(rownames(this_combined)) %in% this_genes)
} else {
  this_genes_index <- which(rownames(this_combined) %in% this_genes)
}


mat <-
  AverageExpression(this_combind, slot = "data")$RNA[this_genes_index, ]
mat <- as.matrix(mat)

#mat <- log1p(mat)
mat <- (mat - as.numeric(rowMeans(mat))) / rowSds(mat)


color_sequence <- seq(min(mat), max(mat), length = 3) / 1.5
color_fun = colorRamp2(color_sequence, c("#3266AD", "white", "#B12424"), space = "RGB")



mat <- mat[this_genes, ]

ht1 <- Heatmap(
  mat,
  show_column_names = T,
  cluster_rows = F,
  show_row_names = T,
  row_km = NULL,
  show_column_dend = F,
  col = color_fun,
  heatmap_legend_param = list(
    title = "Relative\nexpression",
    at = c(min(mat), 0,  max(mat)),
    labels = c("low", "", "high")
  ),
  column_order = 1:ncol(mat),
  column_names_rot = 30
)




pdf(
  paste0("../figure/heatmap_avg_top3_KO.pdf"),
  width = 12,
  height = 8
)

print(draw(ht1, padding = unit(c(0, 30, 0, 0), "mm")))
dev.off()


pdf(
  paste0("../figure/dotplot_avg_top3_KO.pdf"),
  width = 14,
  height = 6
)

DotPlot(this_combined, features = this_genes, cols = two_color) + theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()

```


# Heat by cells selected CT x markers

```{r}

this_combined <- selected_combined
DefaultAssay(this_combined) <- "RNA"
this_combined <- ScaleData(this_combined)
Idents(this_combined) <- this_combined$ct_day

mapal <- colorRampPalette(RColorBrewer::brewer.pal(6,"RdBu"))(128)

pdf(
  paste0("../figure/heatmap_cells_top3.pdf"),
  width = 12,
  height = 6
)

print(DoHeatmap(this_combined, features = this_genes, angle = 90, label = F) + scale_fill_gradientn(colours = rev(mapal))) 
dev.off()

```

# Heat by cells selected CT x markers (only WT)

```{r}

Idents(selected_combined) <- selected_combined$condition
this_combined <- subset(selected_combined, idents = "Con")
Idents(this_combined) <- this_combined$ct_day
DefaultAssay(this_combined) <- "RNA"
this_combined <- ScaleData(this_combined)

#this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combined) <- this_combined$ct_day

mapal <- colorRampPalette(RColorBrewer::brewer.pal(6,"RdBu"))(128)

pdf(
  paste0("../figure/heatmap_cells_top3_WT.pdf"),
  width = 12,
  height = 6
)

print(DoHeatmap(this_combined, features = this_genes, angle = 90, label = F) + scale_fill_gradientn(colours = rev(mapal))) 
dev.off()

```


# Heat by cells selected CT x markers (only ko)

```{r}

Idents(selected_combined) <- selected_combined$condition
this_combined <- subset(selected_combined, idents = "KO")
Idents(this_combined) <- this_combined$ct_day
DefaultAssay(this_combined) <- "RNA"
this_combined <- ScaleData(this_combined)


#this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combined) <- this_combined$ct_day

mapal <- colorRampPalette(RColorBrewer::brewer.pal(6,"RdBu"))(128)

pdf(
  paste0("../figure/heatmap_cells_top3_KO.pdf"),
  width = 12,
  height = 6
)

print(DoHeatmap(this_combined, features = this_genes, angle = 90, label = F) + scale_fill_gradientn(colours = rev(mapal))) 
dev.off()

```



# Archive below

# Heat by avg selected CT x top5

```{r}

this_combined <- selected_combined

DefaultAssay(this_combined) <- "RNA"

#cts_markers <- FindAllMarkers(this_combined, logfc.threshold = 0.6)
this_genes <- cts_markers %>%
  group_by(cluster) %>%
  arrange(cluster) %>%
  top_n(5, avg_log2FC) %>%
  pull(gene) %>%
  unique()

Idents(this_combind) <- as.factor(this_combind$cell_type)

this_genes_index <- which(rownames(this_combined) %in% this_genes)

mat <-
  AverageExpression(this_combind, slot = "data")$RNA[this_genes_index, ]
mat <- as.matrix(mat)

#mat <- log1p(mat)
mat <- (mat - as.numeric(rowMeans(mat))) / rowSds(mat)

mat <- mat[this_genes, ]

color_sequence <- seq(min(mat), max(mat), length = 3) / 1
color_fun = colorRamp2(color_sequence, c("#3266AD", "white", "#B12424"), space = "RGB")


ht1 <- Heatmap(
  mat,
  show_column_names = T,
  cluster_rows = F,
  show_row_names = T,
  row_km = NULL,
  show_column_dend = F,
  col = color_fun,
  heatmap_legend_param = list(
    title = "Relative\nexpression",
    at = c(min(mat), 0,  max(mat)),
    labels = c("low", "", "high")
  ),
  column_order = 1:ncol(mat),
  column_names_rot = 30
)


pdf(
  paste0("../figure/heatmap_avg_selected_ct_top5.pdf"),
  width = 8,
  height = 8
)

print(draw(ht1, padding = unit(c(0, 30, 0, 0), "mm")))
dev.off()

```

# Heat by cells selected CT x top5

```{r}

this_genes <- cts_markers %>%
  group_by(cluster) %>%
  arrange(cluster) %>%
  top_n(5, avg_log2FC) %>%
  pull(gene) %>%
  unique()


this_combined <- selected_combined

DefaultAssay(this_combined) <- "RNA"

#this_combind <- ScaleData(this_combined, verbose = F)
Idents(this_combind) <- as.factor(this_combind$cell_type)

mapal <- colorRampPalette(RColorBrewer::brewer.pal(11,"RdBu"))(256)


pdf(
  paste0("../figure/heatmap_cells_selected_ct_top5.pdf"),
  width = 8,
  height = 10
)

print(DoHeatmap(this_combind, features = this_genes) + scale_fill_gradientn(colours = rev(mapal)))
dev.off()
```


# DEG heatmap on same day, same CT

day 30

"Atrial cardiomyocytes"
"Ventricular cardiomyocytes"

```{r}

this_day <- 'day14'
this_dir <- paste0("../result/deg/", this_day, "/")
this_files <- list.files(this_dir, pattern = "KO_vs_Con.csv")
this_files
this_file <- this_files[1]
for(this_file in this_files) {
  this_combined <-  qs::qread(paste0("../integration/",this_day,"_harmony.qsave"))
  DefaultAssay(this_combined) <- "RNA"
  cts_markers <-
    read.csv(paste0("../result/deg/", this_day, "/",this_file)) 
  
  top_up <- cts_markers %>%
    top_n(20, avg_log2FC) %>%
    pull(gene)
  top_down <- cts_markers %>%
    top_n(-20, avg_log2FC) %>%
    pull(gene)
  
  this_genes <- c(top_up, top_down)
  
  mapal <- colorRampPalette(RColorBrewer::brewer.pal(11, "RdBu"))(256)
  
  pdf(
    paste0(
      "../figure/heatmap_",this_day,"_top20_by_cell_",this_file,".pdf"
    ),
    width = 8,
    height = 10
  )
  
  Idents(this_combined) <- this_combined$orig.ident
  print(
    DoHeatmap(this_combined, features = this_genes) + scale_fill_gradientn(colours = rev(mapal))
  )
  dev.off()
}



```



# Heat by avg selected CT x top20

```{r}

Idents(combined) <- combined$cell_type
this_combind <- subset(combined, idents = c("Atrial cardiomyocytes"))
Idents(this_combind) <- this_combind$orig.ident
DefaultAssay(this_combined) <- "RNA"
#cts_markers <- FindMarkers(this_combind, ident.1 = "N1KO30","Con30")

top_up <- cts_markers %>%
  rownames_to_column("gene") %>%
  top_n(20, avg_log2FC) %>%
  pull(gene)
top_down <- cts_markers %>%
  rownames_to_column("gene") %>%
  top_n(-20, avg_log2FC) %>%
  pull(gene)

this_genes <- c(top_up, top_down)


this_genes_index <- which(rownames(this_combined) %in% this_genes)

mat <-
  AverageExpression(this_combind, slot = "data")$RNA[this_genes_index, ]
mat <- as.matrix(mat)

mat <- log1p(mat)
mat <- (mat - as.numeric(rowMeans(mat)))

mat <- mat[this_genes, ]

color_sequence <- seq(min(mat), max(mat), length = 3) / 1
color_fun = colorRamp2(color_sequence, c("#3266AD", "white", "#B12424"), space = "RGB")


ht1 <- Heatmap(
  mat,
  show_column_names = T,
  cluster_rows = F,
  show_row_names = T,
  row_km = NULL,
  show_column_dend = F,
  col = color_fun,
  heatmap_legend_param = list(
    title = "Relative\nexpression",
    at = c(min(mat), 0,  max(mat)),
    labels = c("low", "", "high")
  ),
  column_order = 1:ncol(mat),
  column_names_rot = 30
)


pdf(
  paste0("../figure/heatmap_day30_Atrial_cardiomyocytes_top20_by_avg.pdf"),
  width = 8,
  height = 8
)

print(draw(ht1, padding = unit(c(0, 30, 0, 0), "mm")))
dev.off()

```

## Volcano

```{r}

p1 <- EnhancedVolcano(
  cts_markers,
  lab = rownames(cts_markers),
  x = 'avg_log2FC',
  y = 'p_val_adj',
  selectLab = c(top_up, top_down),
  pCutoff = 0.05,
  FCcutoff = 0.25,
  drawConnectors = TRUE
)

pdf(
  paste0("../figure/volcano_day30_Atrial_cardiomyocytes_ko_vs_wt.pdf"),
  width = 8,
  height = 8
)
print(p1)
dev.off()


```
